Local Type Inference for Polarised System F with Existentials

### 2023-10-24 oopsla24/tex/01-introduction.tex
foo

However, the universal quantifier is not the only quantifier! Dual to the universal quantifer $\forall$ is the existential quantifier $\exists$. Even though existential types have an equal logical status to universal types, they have been studied much less frequently in the literature. The most widely-used algorithm for handling existentials is the algorithm of Odersky and Laufer. This algorithm treats existentials in a second-class fashion: they are not explicit connectives of the logic, but rather are tied to datatype declarations. As a result, packing and unpacking existential types is tied to constructor introductions and constructor eliminations in pattern matches.

>>>>

However, the universal quantifier is not the only quantifier! Dual to the universal quantifer $\forall$ is the existential quantifier $\exists$. Even though existential types have an equal logical status to universal types, they have been studied much less frequently in the literature. The most widely-used algorithm for handling existentials is the algorithm of Odersky and Laufer. This algorithm treats existentials in a second-class fashion: they are not explicit connectives of the logic, but rather are tied to datatype declarations. As a result, packing and unpacking existential types is tied to constructor introductions and constructor eliminations in pattern matches. This allows Damas-Milner inference to continue to work almost unchanged, but it does come at the cost of losing first-class existentials and also of restricting the witness types to monomorphic types.

There has been a limited amount of work on support for existential types as a first-class concept. In an unpublished draft, \citet{leijen06} studied an extension of Damas-Milner inference in which type schemes contain alternating universal and existential quantifiers. Quantifiers still range over monotypes, and higher-rank polymorphism is not permitted. More recently, \citet{dk19} studed type inference for existential types in the context of GADT type inference, which, while still predicative, supported higher-rank types (i.e., quantifiers can occur anywhere inside of a type scheme). \citet{existential-crisis} propose a system which is intermediate between prenex and higher-rank polymorphism, but which

All of these papers are restricted to \emph{predicative} quantification, whereas existential types in full System F are \emph{impredicative} -- that is, quantifiers can be instantiated with
